---
title: "PA1_Template.Rmd"
author: "mt"
date: "October 1, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(tidyr)
library(readr)
library(ggplot2)

```

Github location is  <https://github.com/mtrinn/RepData_PeerAssessment1>.

## Loading and preprocessing data


We load the data activity:

```{r}
activity=read_csv('activity.csv')

summary(activity)

head(activity,5)

```

We remove the NA observations in step:

```{r}
activitynona <- activity %>% filter(!is.na(steps))
summary(activitynona)
str(activitynona)
```
## What is mean total number of steps taken per day?

We calculate the total number of steps taken per day:

```{r}
ndaysteps <- activitynona %>% group_by(date) %>% summarize(ssteps=sum(steps)) 
head(ndaysteps)
```

We plot a histogram of the total number of steps taken each day. We calculate first the total number of steps for each day:

```{r}
qplot(ndaysteps$ssteps, geom="histogram") 
```

We calculate and report the mean and median of the total number of steps taken per day:

```{r}
nsteps <- activitynona %>% group_by(date) %>% summarize(ssteps=sum(steps)) %>% summarize(mean(ssteps), median(ssteps)) 
nsteps
# answer is
#  `mean(ssteps)` `median(ssteps)`
#           <dbl>            <int>
#       10766.19            10765
```

## What is the average daily activity pattern?

We calculate the average number of steps per 5-minute interval averaged across all days and plot the average steps (y-axis)against the intervals (x-axis):

```{r}
avsteps <- activitynona %>% group_by(interval) %>% summarize(msteps=mean(steps)) 

qplot(avsteps$interval,avsteps$msteps)+geom_line()
```

We look for the interval with the maximum number of steps:

```{r}
avsteps[avsteps[,2]==max(avsteps$msteps),]
# A tibble: 1 x 2
#  interval   msteps
#     <int>    <dbl>
#      835 206.1698
# answer is interval 835
```

## Imputing missing values

We look for the number of missing values, we can use the function summary on the original activity data:

```{r}
summary(activity)
 #Min.   :  0.00   Min.   :2012-10-01   Min.   :   0.0  
 #1st Qu.:  0.00   1st Qu.:2012-10-16   1st Qu.: 588.8  
 #Median :  0.00   Median :2012-10-31   Median :1177.5  
 #Mean   : 37.38   Mean   :2012-10-31   Mean   :1177.5  
 #3rd Qu.: 12.00   3rd Qu.:2012-11-15   3rd Qu.:1766.2  
 #Max.   :806.00   Max.   :2012-11-30   Max.   :2355.0  
 #NA's   :2304  
# answer: there are 2,304 missing values
```

We can replace missing values by the average number of steps for the interval:

```{r}
activity1 <- activity %>% group_by(interval) %>% mutate(steps=ifelse(is.na(steps),mean(steps,na.rm=TRUE),steps))
head(activity1,10)
```

We make a histogram of the total number of steps taken each day:

```{r}
ndaysteps1 <- activity1 %>% group_by(date) %>% summarize(ssteps=sum(steps)) 

qplot(ndaysteps1$ssteps, geom="histogram") 
```

We calculate and report the mean and median total number of steps taken per day.

```{r}
nsteps1 <- activity1 %>% group_by(date) %>% summarize(ssteps=sum(steps)) %>% summarize(mean(ssteps), median(ssteps)) 
nsteps1
# answer:
#`mean(ssteps)` `median(ssteps)`
#           <dbl>            <dbl>
#       10766.19         10766.19
  
```
We see that the estimates of the total number of steps is the same for the mean (10766.19) and slighlty higher for the median (10766.16 vs 10765 before).

## Are there differences in activity patterns between weekdays and weekends?

We create weekday column:

```{r}
activity2 <- activity1  %>% mutate(day=ifelse((weekdays(date,abbreviate=TRUE)=="Sat" | weekdays(date,abbreviate=TRUE)=="sun"),"weekend","weekday"))
```


```{r}
avsteps2 <- activity2 %>% group_by(interval,day) %>% summarize(msteps=mean(steps)) 

p<-ggplot(avsteps2,aes(interval,msteps))+geom_line()
p + facet_grid(. ~ day)

```

We see during the week there is lot activities in the morning, probably before people go to work. During the weekend, there is also acitivities in the afternoon.
